The Host Galaxies of 106 Rapidly Evolving Transients discovered by the Dark Energy Survey

### 2020-07-22 DES_RET_Hosts_replyref.tex
Updates from Overleaf

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a recently-discovered group of astrophysical events that display rapid luminosity evolution. RETs typically rise to peak in less than 10 days and fade within 30, a timescale unlikely to be compatible with the decay of Nickel-56 that drives conventional SNe. Their peak luminosity spans a range of $-15 <M_g<-22.5$, with some events observed at redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, and thus their origin and explosion/emission mechanism remains unexplained. In this paper we define the largest sample of RETs to date, comprising 106 objects from the Dark Energy Survey, and perform the most comprehensive analysis of RET host galaxies. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities ([O/H]) for 37. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs ($\langle\log \left(\mathrm{sSFR}\right)\rangle \sim -9.6$ ), indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies $\langle12+\log\left(\mathrm{O/H}\right) \rangle \sim 9.4$. There are no clear relationships between properties of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

Recently, inspection of high-cadence, large-area survey data sets have revealed more exotic transients that are difficult to explain with conventional models. \citet{Drout2014} presented a sample of rapidly evolving transients (RETs; also termed `Fast Blue Optical Transients' - FBOTs or `Fast Evolving Luminous Transients' - FELTs) in the Pan-STARRS survey (PS1). \citet{Arcavi2016} reported three rapidly rising, highly luminous objects in the Supernova Legacy Survey (SNLS) and one in the Palomar Transient Factory (PTF), of which one (SNLS04D4ec) declines rapidly. \citet[hereafter P18]{Pursiainen2018} expanded the known number of RETs to beyond 80 with their sample from the Dark Energy Survey (DES), spanning a redshift range of $\sim 0$ to $>1$. A further sample of five objects has been discovered by the Hyper Suprime-Cam Subaru Strategic Program (SSP) Transient Survey \citep{Tampo2020}. \replyref{Although the search and selection criteria vary between the above samples, the objects} typically rise to peak brightness in less than 10 days, and decline to 10\% of their peak brightness within 30 days, much faster than typical SNe. The photometric measurements of the PS1 and DES RETs seem to be well described by expanding blackbodies, although a handful show declining photospheric radii from the first detection. Due to the rapid nature of their lightcurves and location at high-redshift, spectral coverage is sparse and signal-to-noise ratio (SNR) is low, such that there has not yet been a conclusive detection of absorption or emission features from the transients and thus the physical mechanism responsible for their rapid evolution remains unexplained.

There are a limited number of events detected in the local Universe whose properties appear consistent with the RETs seen at cosmological distances in the PS1 and DES samples, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. The transient declined from its discovery, with constraints on the rise time of 1 day, and from X-rays through to radio wavelengths did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}) . There are many diverse explanations for the power source of AT2018cow \replyref{proposed} in the literature, including: \replyref{engine-driven stellar explosions \citep{Ho2019,Margutti2019,Mohan2020}}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Piro2020,Uno2020}.
Other nearby rapid transients \replyref{that would pass the DES search criteria} include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the accretion-induced collapse of a white dwarf or a white dwarf-neutron star merger, KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material, \replyref{SN2018gep \citep{Ho2019a} which is a rapidly-declining, nearby SN Ic-BL, and the H-rich, relativistic CSS161010 \citep{Coppejans2020}. At higher redshift there is one RET -- ZTF18abvkwla \citep{Ho2020}.}

In this paper, we present the first comprehensive study of the host galaxies of RETs. We make use of the final DES sample, which builds on \citetalias{Pursiainen2018} by adding the 5th and final season of DES-SN observations as well as more refined selection techniques. Using the deep DES photometry from \citet[hereafter W20]{Wiseman2020} and spectra from OzDES \citep{Lidman2020} we derive host galaxy properties and compare them to samples of host galaxies of CCSNe, LGRBs, and SLSNe, as well as the individual local rapid transients. For clarity, we will use the term RET to refer only to events in the high-redshift samples of DES and PS1.

@@

Transient name &  (J2000  & (J2000)  &  (AB)    &         &         &   (hours)         &               \\

@@

\section{Sample selection}
\label{sec:sample}
We derive our sample from the 106 RETs discovered in the 5-year DES-SN transient survey. This number expands upon the 72 of \citetalias{Pursiainen2018}. The first reason for the increased sample size is the use of the 5th year of DES-SN, as \citetalias{Pursiainen2018} were only able to make use of the first four years. By imposing the \citetalias{Pursiainen2018} selection criteria on season 5, the sample is increased to 92 objects. The second reason is an update to the sample selection technique, outlined in the following subsection, which adds a further 14 transients.

@@

The Dark Energy Survey (DES) makes use of the Dark Energy Camera (DECam; \citealt{Flaugher2015}) to survey $5000\deg^2$ of the southern sky between 2013 and 2019. The supernova programme (DES-SN), designed primarily to detect type Ia SNe (SNe Ia) for cosmological measurements \citep{Bernstein2012}, consisted of five 6-month seasons of approximately 7 day cadence in ten single pointings, known as the SN fields which cover a total area of $\sim27 \deg^2$. Of these fields, two were observed with longer exposure times and are referred to as deep fields with remaining eight known as shallow fields; the resulting approximate single-visit depths are 23.5 mag (shallow) and 24.5 mag (deep) in the DES $r$-band. Transient detection was performed by difference imaging using a custom pipeline \citep{Kessler2015}. Transient candidates were vetted via machine-learning techniques \citep{Goldstein2015}, leading to $\sim 30,000$ viable supernova candidates over the full five seasons. DES-SN included an extensive spectroscopic follow-up program to identify transients and measure spectroscopic redshifts of host galaxies \citep{DAndrea2018}. A full description of the search for rapid transients in DES can be found in \citetalias{Pursiainen2018}.

@@

The original method of finding RETs in the DES-SN data (and presented in \citetalias{Pursiainen2018}) was designed to be simple and used light curve modelling with Gaussian and linear fits. The simplistic method made it possible to look for exotic transients without knowing their observed characteristics beforehand and resulted in a large sample of photometrically selected fast transients. However, as the search was simplistic and relied heavily on visual inspection of the available data (etc. images, light curves, host galaxy information), it is impossible to quantify the completeness of the sample. For instance, due to the large redshift range within the sample it is entirely possible that distant events with longer rise times could have been missed due to time dilation stretching their lightcurves, while faster events and low redshift may evolve on timescales quicker than the survey cadence. Here, a more sophisticated search method is presented. As only a fraction of transients in DES-SN have redshift information from their host galaxies, we perform the search in the observer frame. The key features of RETs that separate them from most traditional SNe types are the fast light curve evolution (rise to peak in $\lesssim15$ days) and blue colour at peak ($g-r \lesssim 0.5$). Even though both of these quantities depend on the redshift of the transient, they still effectively distinguish the the fast events from traditional SN types. We thus attempt to select a sample based on observed rise times and colours at peak brightness.

@@

\subsubsection{Photometric definition of RETs \label{subsubsec:trise_gr}}
To make an improved selection of RETs we use a parameter space described by observed $g-r$ colour and the time taken to rise from non-detection to peak $r$-band magnitude in the observer frame. Using rise times and colours from GP lightcurves, we populate this parameter space with the sample of 72 RETs from the \citetalias{Pursiainen2018} method, updated with 20 extra objects found using that method in the fifth season of DES. We add spectroscopically confirmed SNe of types Ia, Ibc, II, and SLSNe observed by DES in order to verify that they are rejected by the search method. We keep objects passing the following selection requirements (cuts):
\begin{enumerate}
\item The transient was detected in only one DES-SN observing season.
\item Maximum observed brightness in both $g$- and $r$-bands was brighter than 24 mag (in the eight `shallow' DES-SN fields) or 25 mag (in the two `deep' fields), as in \citetalias{Pursiainen2018}.
%\item $g$- and $r$-band observations used for the colour had to be taken within 2 days of each other in the observer frame.
\end{enumerate}

@@

In order to apply selection criterion i) from Section \ref{subsubsec:trise_gr} it is necessary to distinguish between DES-SN candidates that are truly multi-season events (typically active galactic nuclei; AGN) and those that are single-season events with spurious detections in other seasons. 
To detect and filter out AGN we use a basic convolutional neural network (CNN) classifier. We train the CNN on spectroscopically confirmed SNe of all types as well as the 92 RETs identified using the \citetalias{Pursiainen2018} method (181 objects) and on spectroscopically typed AGN (182 objects), and use it to separate the sample into two photometric subtypes: AGN-like and SN-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set (391 SNe and 79 AGN). No SNe were classified as AGN-like. The remaining AGN classified as SNe-like are removed by manual vetting later in the process. The CNN does not separate the SN-like objects into RETs and other SN subtypes: this is done in the subsequent processing step.

\subsubsection{Final DES RET sample}
The GP lightcurves of all DES-SN candidates are classified as AGN-like or SNe-like by the CNN, and are then subject to the lightcurve quality cuts i) and ii) (Section \ref{subsubsec:trise_gr}), resulting in 2259 objects, of which 939 lie inside the colour and rise-time region which we defined for RETs. These objects are subject to a further set of cuts in order to remove remaining contaminants. We impose a cut based on a SN classifier (\texttt{PSNID}; \citealt{Sako2008}), that returns a normalised goodness of fit to different SN Ia and CCSN templates, along with a Bayesian probability of it being a SN Ia. To remove highly-probable SNe Ia, we use threshold probabilities of P(Ia)$<0.91$ and P(Ia; Bayes)$<0.82$ respectively to the above algorithms, which removes 46 objects from the RET parameter space. In order to further remove longer-lived SNe, the decline time to half of the peak brightness must be $<24$ days. This removes 347 SNe, resulting in 546 objects remaining inside the parameter space. The final 546 transients have been visually inspected, with the majority rejected for clearly being spurious detections, obvious multi-season variability that was not picked up by the CNN, or showing a longer timescale decline.

Using the above method recovers 87 of the 92 RETs found using the \citetalias{Pursiainen2018} technique, and adds a further 14. The five were not recovered as their GP lightcurves were fainter than the limits given above in either g or r band.
We refer to the resulting sample as DES RETs. Of the 106 objects in the sample, 96 have a host galaxy detected in deep host galaxy photometry of \citetalias{Wiseman2020} when using the Directional Light Radius method \citep{Sullivan2006} to associate hosts as per \citetalias{Wiseman2020}. Of these, 49 have a host galaxy spectroscopic redshift which we access through an internal release of the OzDES Global Redshift Catalog (GRC; v.2020\_01\_04). The full OzDES redshift catalogue will be available alongside the public data release detailed in \citet{Lidman2020}. A further three have redshifts obtained from narrow lines observed in spectra of the transients themselves. We do not consider these three objects for the analysis, since we are unable to separate transient and host contributions to the spectra.
A selection of the host galaxies is shown in Fig.~\ref{fig:mosaic}, centred on the location of the transient. The figure showcases the diversity of host galaxy morphologies and colours, while also displaying the limitations of ground-based observations of high-redshift, relatively small galaxies in terms of spatial resolution.
Fig.~\ref{fig:z_dist} shows the distribution of redshifts amongst the 49 hosts for which such a measurement was possible. The effect that the redshift selection function has on the results is discussed in Section \ref{subsec:disc_bias}.
The observational properties of the 96 detected hosts are displayed in Table \ref{tab:obs}. We highlight objects that were not presented in \citetalias{Pursiainen2018}. We also highlight a small subset of objects for which the redshift has taken on a new value than that presented in \citetalias{Pursiainen2018} due to further OzDES observations leading to a more accurate determination.

@@

\subsubsection{RETs \label{subsubsec:compare_rets}}
Since the DES sample of RETs is by far the largest discovered to date, there is no other large sample of RETs with which to compare host galaxy properties. \citet{Drout2014} present host galaxies of 10 RETs discovered in the Pan-STARRS survey, with measurements of stellar masses, SFRs, and metallicities. We also compare with the host galaxy of SNLS04D4ec \citep{Arcavi2016}. To this we add the low-redshift transient AT2018cow (nicknamed ``The Cow''). The host galaxy of AT2018cow has been studied with photometric measurements \citep{Perley2019} as well as with an integral field spectrograph \citep{Lyman2020}, with consistent results. For our comparison, we use the galaxy-averaged stellar mass and SFR from \citet{Lyman2020} and the metallicity from the host nucleus as reported by both \citet{Morokuma-Matsui2019} and \citet{Lyman2020} as it best represents the method of obtaining spectra for our RET sample. We further compare to SN2018gep with data from  \citet{Ho2019}, and ZTF18abvkwla (nicknamed ``The Koala'') from \citet{Ho2020}.

\subsubsection{SNe and GRBs \label{subsubsec:compare_CCSNe}}

In compiling a set of comparison samples, we aim for the least biased selections possible. This requires surveys to be untargetted (they were not monitoring certain galaxies in order to search for SNe), ideally complete, and also covering a similar redshift range to the RETs. While in practice the second and third of these criteria are difficult to achieve, particularly with the fainter CCSNe, we are able to choose comparison samples from untargetted surveys to mitigate initial selection biases.

To compare with CCSNe, we draw on the untargetted sample of 47 SNe II from the Palomar Transient Factory (PTF; \citealt{Stoll2013}), which is likely complete in terms of hosts (all SNe have an associated host). While this sample lies at much lower redshift than the DES RETs (a maximum of 0.18 and mean of 0.05), redshift evolution is easier to account for in a less biased way than correcting for unknown incompleteness. We add to this the compilation of 56 untargeted SESNe from \citet{Sanders2012}, with a maximum redshift of 0.26 and a mean of 0.05. Since \citet{Sanders2012} do not report host galaxy magnitudes, stellar masses or SFRs, we cross-match the SN positions with the the Sloan Digital Sky Survey (SDSS; \citealt{York2000}) Data Release 16 (DR16; \citealt{Ahumada2019}) and perform our own SED fit using the method outlined in Section \ref{subsec:sedfit}. We are able to do this SED fit for 38 objects, with the others lying outside of the SDSS footprint.

@@

To show how RETs compare to the galaxy population as a whole, we use a sample of $\sim800,000$ measurements from the MPA-JHU catalogues of stellar masses (based on the methods of \citealt{Kauffmann2003,Salim2007}), SFRs (based off \citealt{Brinchmann2004}), and metallicities (based off \citealt{Tremonti2004}) from the  catalogues of SDSS Data Relase 7 (DR7; \citealt{Abazajian2009}). The mean redshift is 0.08 which is much lower than for the RETs, such that significant evolution in the galaxy population has happened between the majority of RET hosts and the SDSS sample. We do not correct for this, but take it into account when analysing our findings. 
\section{Host galaxy observations}
\label{sec:obs}
\subsection{Photometry \label{subsec:phot}}

>>>>

% Abstract of the paper
\begin{abstract}
Rapidly evolving transients (RETs), also termed fast blue optical transients (FBOTs), are a recently-discovered group of astrophysical events that display rapid luminosity evolution. RETs typically rise to peak in less than 10 days and fade within 30, a timescale unlikely to be compatible with the decay of Nickel-56 that drives conventional SNe. Their peak luminosity spans a range of $-15 <M_g<-22.5$, with some events observed at redshifts greater than 1. Their evolution on fast timescales has hindered high quality follow-up observations, and thus their origin and explosion/emission mechanism remains unexplained. In this paper we define the largest sample of RETs to date, comprising 106 objects from the Dark Energy Survey, and perform the most comprehensive analysis of RET host galaxies. Using deep-stacked photometry and emission-lines from OzDES spectroscopy, we derive stellar masses and star-formation rates (SFRs) for 49 host galaxies, and metallicities ([O/H]) for 37. We find that RETs explode exclusively in star-forming galaxies and are thus likely associated with massive stars. Comparing RET hosts to samples of host galaxies of other explosive transients as well as field galaxies, we find that RETs prefer galaxies with high specific SFRs ($\langle\log \left(\mathrm{sSFR}\right)\rangle \sim -9.6$ ), indicating a link to young stellar populations, similar to stripped-envelope SNe. RET hosts appear to show a lack of chemical enrichment, their metallicities akin to long duration gamma-ray bursts and superluminous SN host galaxies $\langle12+\log\left(\mathrm{O/H}\right) \rangle \sim 9.4$. There are no clear relationships between \replyref{mass or SFR} of the host galaxies and the peak magnitudes or decline rates of the transients themselves.

@@

Recently, inspection of high-cadence, large-area survey data sets have revealed more exotic transients that are difficult to explain with conventional models. \citet{Drout2014} presented a sample of rapidly evolving transients (RETs; also termed `Fast Blue Optical Transients' - FBOTs or `Fast Evolving Luminous Transients' - FELTs) in the Pan-STARRS survey (PS1), \replyref{defined as having time above half-brightness ($t_{\mathrm{half}}) \leq12$d}. \citet{Arcavi2016} reported three rapidly rising, highly luminous objects in the Supernova Legacy Survey (SNLS) and one in the Palomar Transient Factory (PTF), of which one (SNLS04D4ec) declines rapidly. \citet[hereafter P18]{Pursiainen2018} expanded the known number of RETs to beyond 80 with their sample from the Dark Energy Survey (DES), spanning a redshift range of $\sim 0$ to $>1$. A further sample of five objects has been discovered by the Hyper Suprime-Cam Subaru Strategic Program (SSP) Transient Survey \citep{Tampo2020}. \replyref{Although the search and selection criteria vary between the above samples, the objects} typically rise to peak brightness in less than 10 days, and decline to 10\% of their peak brightness within 30 days, much faster than typical SNe. The photometric measurements of the PS1 and DES RETs seem to be well described by expanding blackbodies, although a handful show declining photospheric radii from the first detection. Due to the rapid nature of their lightcurves and location at high-redshift, spectral coverage is sparse and signal-to-noise ratio (SNR) is low, such that there has not yet been a conclusive detection of absorption or emission features from the transients and thus the physical mechanism responsible for their rapid evolution remains unexplained.

There are a small number of other objects in the literature that would pass the DES RET criteria, including SNLS04D4ec \citep{Arcavi2016} at $z\sim0.59$, ZTF18abvkwla \replyref{whose rapid evolution has been attributed to shock interaction, while radio emission can be explained by a collimated jet \citep{Ho2020} at $z\sim0.27$, and the rapidly-evolving SN Ic-BL-like iPTF16asu \citep{Whitesides2017} at $z\sim0.19$.}

There are also a limited number of events detected in the local Universe whose properties appear consistent with the RETs seen at cosmological distances in the PS1 and DES samples, the most widely studied of which is AT2018cow \citep[e.g.][]{Prentice2018,Margutti2019,Perley2019}. \replyref{AT2018cow has drawn many comparisons to the cosmological RETs from DES and PS1 \citep[e.g.][]{Perley2019,Margutti2019,Fox2019,Mohan2020} due to its rapid evolution and blue colour. AT2018cow declined from its discovery, with constraints on the rise time of 1 day, and from X-rays through to radio wavelengths did not resemble any known SN, GRB afterglow, or kilonova (KN; \citealt{Ho2019}). It displayed a contracting photosphere as well as evidence for central-engine power alongside an unusual spectrum that showed similarities to broad-lined SNe Ic (SN Ic-bl) at early stages \citep[e.g.][]{Xu2018,Izzo2018}, developing to something entirely different at later epochs \citep{Perley2019} with hints of similarities to interacting SNe Ibn \citep{Fox2019}.} There are many diverse explanations for the power source of AT2018cow \replyref{proposed} in the literature, including: \replyref{engine-driven stellar explosions \citep{Ho2019,Margutti2019,Mohan2020}}; electron capture collapse of merged white dwarfs \citep{Lyutikov2019}; a tidal disruption event (TDE) of a white dwarf \citep{Kuin2019} or of a main sequence star by an intermediate mass black hole \citep{Perley2019}; common envelope jets supernova (CEJSN; \citealt{Soker2019}); or a wind-driven transient \citep{Piro2020,Uno2020}.
Other nearby \replyref{($z<0.1$)} rapid transients \replyref{that would pass the DES search criteria} include the local fast-declining SN-like transient SN2018kzr \citep{McBrien2019} which is explained by the accretion-induced collapse of a white dwarf or a white dwarf-neutron star merger, KSN-2015K \citep{Rest2018} whose fast rise and decline is explained by the shock of an SN running into previously-expelled material, \replyref{SN2018gep \citep{Ho2019a} \replyref{which is a rapidly-declining, spectroscopically classified, nearby SN Ic-BL with a rapid rise to peak}, and the H-rich, relativistic CSS161010 \citep{Coppejans2020}.}

In this paper, we present the first comprehensive study of the host galaxies of RETs. We make use of the final DES sample, which builds on \citetalias{Pursiainen2018} by adding the 5th and final season of DES-SN observations as well as more refined selection techniques. Using the deep DES photometry from \citet[hereafter W20]{Wiseman2020} and spectra from OzDES \citep{Lidman2020} we derive host galaxy properties and compare them to samples of host galaxies of CCSNe, LGRBs, and SLSNe, as well as the individual local rapid transients.

@@

Transient name &  (\replyref{deg }J2000)  & (\replyref{deg} J2000)  &  (AB)    &         &         &   (hours)         &               \\

@@

\section{Sample selection}
\label{sec:sample}
\replyref{Our sample of RETs comprises 106 events} discovered in the 5-year DES-SN transient survey. This number expands upon the 72 of \citetalias{Pursiainen2018}. The first reason for the increased sample size is the use of the 5th year of DES-SN, as \citetalias{Pursiainen2018} were only able to make use of the first four years. By imposing the \citetalias{Pursiainen2018} selection criteria on season 5, the sample is increased to 92 objects. The second reason is an update to the sample selection technique, outlined in the following subsection, which adds a further 14 transients.

@@

The Dark Energy Survey (DES) makes use of the Dark Energy Camera (DECam; \citealt{Flaugher2015}) to survey $5000\deg^2$ of the southern sky between 2013 and 2019. The supernova programme (DES-SN), designed primarily to detect type Ia SNe (SNe Ia) for cosmological measurements \citep{Bernstein2012}, consisted of five 6-month seasons of approximately 7 day cadence in ten single pointings, known as the SN fields which cover a total area of $\sim27 \deg^2$. Of these fields, two were observed with longer exposure times and are referred to as deep fields with \replyref{the} remaining eight known as shallow fields; the resulting approximate single-visit depths are 23.5 mag (shallow) and 24.5 mag (deep) in the DES $r$-band. Transient detection was performed by difference imaging using a custom pipeline \citep{Kessler2015}. Transient candidates were vetted via machine-learning techniques \citep{Goldstein2015}, leading to $\sim 30,000$ viable supernova candidates over the full five seasons. DES-SN included an extensive spectroscopic follow-up program to identify transients and measure spectroscopic redshifts of host galaxies \citep{DAndrea2018}. A full description of the search for rapid transients in DES can be found in \citetalias{Pursiainen2018}.

@@

The original method of finding RETs in the DES-SN data (and presented in \citetalias{Pursiainen2018}) was designed to be simple and used light curve modelling with Gaussian and linear fits. The simplistic method made it possible to look for exotic transients without knowing their observed characteristics beforehand and resulted in a large sample of photometrically selected fast transients. However, as the search was simplistic and relied heavily on visual inspection of the available data (etc. images, light curves, host galaxy information), it is impossible to quantify the completeness of the sample. For instance, due to the large redshift range within the sample it is entirely possible that distant events with longer rise times could have been missed due to time dilation stretching their lightcurves, while faster events \replyref{at} low redshift may evolve on timescales quicker than the survey cadence. Here, a more sophisticated search method is presented. As only a fraction of transients in DES-SN have redshift information from their host galaxies, we perform the search in the observer frame. The key features of RETs that separate them from most traditional SNe types are the fast light curve evolution \replyref{($t_{\mathrm{half}}\leq12$d)} and blue colour at peak ($g-r \lesssim 0.5$). Even though both of these quantities depend on the redshift of the transient \replyref{(timescales, for example, are dilated by a factor of 2 at $z=1$)}, they still effectively distinguish the fast events from traditional SN types. \replyref{Here, we} attempt to select a sample based on observed rise times and colours at peak brightness.

@@

\subsubsection{Photometric definition of RETs \label{subsubsec:trise_gr}}
To make an improved selection of RETs we use a parameter space described by observed $g-r$ colour and the time taken to rise from non-detection to peak $r$-band magnitude in the observer frame. \replyref{While limiting magnitudes are variable, the uncertainty this introduces onto the rise time is minimal compared to that caused by the fact that the rise time is shorter than the cadence of the survey (i.e. $<7$ d) for many RETs, such that the first detection is the peak.} Using rise times and colours from GP lightcurves, we populate this parameter space with the sample of 72 RETs from the \citetalias{Pursiainen2018} method, updated with 20 extra objects found using that method in the fifth season of DES. We add spectroscopically confirmed SNe of types Ia, Ibc, II, and SLSNe observed by DES in order to verify that they are rejected by the search method. We keep objects passing the following selection requirements (cuts):
\begin{enumerate}
\item The transient was detected in only one DES-SN observing season.
\item Maximum observed brightness in both $g$- and $r$-bands was brighter than 24 mag (in the eight `shallow' DES-SN fields) or 25 mag (in the two `deep' fields), as in \citetalias{Pursiainen2018}.
%\item $g$- and $r$-band observations used for the colour had to be taken within 2 days of each other in the observer frame.
\end{enumerate}

\replyref{Light curves are not corrected for Galactic extinction, but this is negligible at $\mathrm{E}(B-V)<<0.1$ in the DES-SN fields.}

@@

In order to apply selection criterion i) from Section \ref{subsubsec:trise_gr} it is necessary to distinguish between DES-SN candidates that are truly multi-season events (typically active galactic nuclei; AGN) and those that are single-season events with spurious detections in other seasons. 
To detect and filter out AGN we use a basic convolutional neural network (CNN) classifier. \replyref{The CNN uses features of the GP light curves in a set of layers in order to ``learn" the difference between SN and AGN lightcurves. We implement two 1-dimensional feature layers: the first searches for evolution over 2-day timescales in all four filters, while the second deals with long-term evolution on 20-day timescales. We include a dropout layer after each filter layer to prevent over-fitting, and a max-pooling layer to increase speed. At the end, a softmax layer outputs a probability that the individual transient is SN-like, for which we use 50\% as the threshold}.  We train the CNN on spectroscopically confirmed SNe of all types as well as the 92 RETs identified using the \citetalias{Pursiainen2018} method (181 objects) and on spectroscopically typed AGN (182 objects), and use it to separate the sample into two photometric subtypes: AGN-like and SN-like. The classifier returns SNe-like objects with an accuracy of 0.992 on the test set (391 SNe and 79 AGN). No SNe were classified as AGN-like. The remaining AGN classified as SNe-like are removed by manual vetting later in the process. The CNN does not separate the SN-like objects into RETs and other SN subtypes: this is done in the subsequent processing step.

\subsubsection{Final DES RET sample}
The GP lightcurves of all DES-SN candidates are classified as AGN-like or SNe-like by the CNN, and are then subject to the lightcurve quality cuts i) and ii) (Section \ref{subsubsec:trise_gr}), resulting in 2259 objects, of which 939 lie inside the colour and rise-time region which we defined for RETs. These objects are subject to a further set of cuts in order to remove remaining contaminants. We impose a cut based on a SN \replyref{light curve} classifier (\texttt{PSNID}; \citealt{Sako2008}), that returns a normalised goodness of fit to different SN Ia and CCSN templates, along with a Bayesian probability of it being a SN Ia. To remove highly-probable SNe Ia, we use threshold probabilities of P(Ia)$<0.91$ and P(Ia; Bayes)$<0.82$ respectively to the above algorithms, which removes 46 objects from the RET parameter space. In order to further remove longer-lived SNe, the decline time to half of the peak brightness must be $<24$ days. This removes 347 SNe, resulting in 546 objects remaining inside the parameter space. The final 546 transients have been visually inspected, with the majority rejected for clearly being spurious detections, obvious multi-season variability that was not picked up by the CNN, or showing a longer timescale decline.

Using the above method recovers 87 of the 92 RETs found using the \citetalias{Pursiainen2018} technique, and adds a further 14. The five were not recovered as their GP lightcurves were fainter than the limits given above in either g or r band.
We refer to the resulting sample as DES RETs. Of the 106 objects in the sample, 96 have a host galaxy detected in deep host galaxy photometry of \citetalias{Wiseman2020} when using the Directional Light Radius method \citep{Sullivan2006} to associate hosts as per \citetalias{Wiseman2020}. Of these, 49 have a host galaxy spectroscopic redshift which we access through an internal release of the OzDES Global Redshift Catalog (GRC; v.2020\_01\_04). The full OzDES redshift catalogue will be available alongside the public data release detailed in \citet{Lidman2020}. A further three have redshifts obtained from narrow lines observed in spectra of the transients themselves. We do not consider these three objects for the analysis, since we are unable to separate transient and host contributions to the spectra.
A selection of the host galaxies is shown in Fig.~\ref{fig:mosaic}, centred on the location of the transient. The figure showcases the diversity of host galaxy morphologies and colours, while also displaying the limitations of ground-based observations of high-redshift, relatively small galaxies in terms of spatial resolution.
Fig.~\ref{fig:z_dist} shows the distribution of redshifts amongst the 49 hosts for which such a measurement was possible. The effect that the redshift selection function has on the results is discussed in Section \ref{subsec:disc_bias}.
The observational properties of the 96 detected hosts are displayed in Table \ref{tab:obs}. We highlight objects that were not presented in \citetalias{Pursiainen2018}. We also highlight a small subset of objects for which \replyref{further OzDES observations have led to a more accurate redshift than that presented in  \citetalias{Pursiainen2018}.}

@@

\subsubsection{RETs \label{subsubsec:compare_rets}}
Since the DES sample of RETs is by far the largest discovered to date, there is no other large sample of RETs with which to compare host galaxy properties. \citet{Drout2014} present host galaxies of 10 RETs discovered in the Pan-STARRS survey, with measurements of stellar masses, SFRs, and metallicities. We also compare with the host galaxy of SNLS04D4ec \citep{Arcavi2016} \replyref{, ZTF18abvkwla (nicknamed ``The Koala'') from \citet{Ho2020}, and iPTF16asu \citep{Whitesides2017}.} To this we add the low-redshift transient AT2018cow (nicknamed ``The Cow''). The host galaxy of AT2018cow has been studied with photometric measurements \citep{Perley2019} as well as with an integral field spectrograph \citep{Lyman2020}, with consistent results. For our comparison, we use the galaxy-averaged stellar mass and SFR from \citet{Lyman2020} and the metallicity from the host nucleus as reported by both \citet{Morokuma-Matsui2019} and \citet{Lyman2020} as it best represents the method of obtaining spectra for our RET sample. We further compare to low-redshift SN2018gep with data from \citet{Ho2019a} \replyref{and CSS161010 with data from \citet{Coppejans2020}.}

\subsubsection{SNe and GRBs \label{subsubsec:compare_CCSNe}}
\replyref{It is likely that the progenitor and explosion scenario of the DES RET sample is heterogeneous in origin and may include some known classes of transient, and as such we choose various samples of host galaxies from the literature to compare the DES RET hosts to.}
In compiling a set of comparison samples, we aim for the least biased selections possible. This requires surveys to be untargetted (they were not monitoring certain galaxies in order to search for SNe), ideally complete, and also covering a similar redshift range to the RETs. While in practice the second and third of these criteria are difficult to achieve, particularly with the fainter CCSNe, we are able to choose comparison samples from untargetted surveys to mitigate initial selection biases.
\replyref{We choose large enough samples that a significant statistical comparison can be drawn, and thus do not split the major groups of SNe into sub-types.}

To compare with CCSNe, we draw on the untargetted sample of 47 SNe II from the Palomar Transient Factory (PTF; \citealt{Stoll2013}), which is likely complete in terms of hosts (all SNe have an associated host). While this sample lies at much lower redshift than the DES RETs (a maximum of 0.18 and mean of 0.05), redshift evolution is easier to account for in a less biased way than correcting for unknown incompleteness. We add to this the compilation of 56 untargeted SESNe from \citet{Sanders2012}, with a maximum redshift of 0.26 and a mean of 0.05. Since \citet{Sanders2012} do not report host galaxy magnitudes, stellar masses or SFRs, we cross-match the SN positions with the Sloan Digital Sky Survey (SDSS; \citealt{York2000}) Data Release 16 (DR16; \citealt{Ahumada2019}) and perform our own SED fit using the method outlined in Section \ref{subsec:sedfit}. We are able to do this SED fit for 38 objects, with the others lying outside of the SDSS footprint.

@@

To show how RETs compare to the galaxy population as a whole, we use a sample of $\sim800,000$ measurements from the MPA-JHU catalogues of stellar masses (based on the methods of \citealt{Kauffmann2003,Salim2007}), SFRs (based off \citealt{Brinchmann2004}), and metallicities (based off \citealt{Tremonti2004}) from the  catalogues of SDSS Data Relase 7 (DR7; \citealt{Abazajian2009}). The mean redshift is 0.08 which is much lower than for the RETs, such that significant evolution in the galaxy population has happened between the majority of RET hosts and the SDSS sample. We do not correct for this, but take it into account when analysing our findings. \replyref{The magnitude-limited nature of the SDSS redshift catalogue means that it is biased against low-mass galaxies, and metallicity estimates requiring strong lines to pass SNR thresholds mean that they too will be biased. Many works exploring extreme transients such as LGRBs, GRB-SNe, and SLSNe, use a complete, volume limited sample of field galaxies such as the Local Volume Legacy Survey (LVL) sample as a comparison \citep[e.g.][]{Chen2017,Modjaz2019,Taggart2019}. However the spectroscopic follow-up of DES RETs was performed as part of the magnitude-limited OzDES survey (Section \ref{subsec:spec}), which has similar selection biases and incompleteness effects to SDSS, albeit shifted to a higher redshift. In comparison, the many dwarf galaxies in LVL would not be detected in DES photometry, or get a redshift measurement from OzDES, and thus such a comparison sample is not representative of the way the DES RET host sample was observed.}

\section{Host galaxy observations}
\label{sec:obs}
\subsection{Photometry \label{subsec:phot}}
